Davos: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-09-13 paper/main.tex
cleaned up tex file formatting

%Thus, rather than constructing fully separate Python environments from scratch, Davos projects work by providing any additions or modifications to the user's existing package environment needed to satisfy their corresponding notebooks' dependencies.
Thus, rather than constructing fully separate Python environments from scratch, Davos projects work by supplementing the user's existing environment with any additional packages (or specific package versions) needed to satisfy 
%their corresponding notebooks' dependencies. 
the dependencies of their corresponding notebooks.
In some cases, this might include every package smuggled into a notebook (e.g., if the notebook is run inside a freshly created, empty virtual environment). In other cases, the user's environment may already contain all required packages, and the notebook's project directory will be unused (in which case it will be deleted automatically when the notebook kernel is shut down). 
But regardless of the extent to which the existing environment is augmented, Davos's project system ensures that all smuggled packages are installed locally and 
%loaded into the notebook successfully, 
can be loaded successfully at runtime,
but that the contents of the user's Python environment are never altered. Additionally, because \texttt{smuggle} statements in a given notebook are evaluated every time it is run, this system also ensures that the notebook's requirements will continue to be satisfied even if the user's Python environment changes. For example, suppose a user has \texttt{NumPy}~\cite{HarrEtal20} v1.24.3 installed in their current Python environment and runs a Davos-enhanced notebook that smuggles \texttt{NumPy} with ``\texttt{numpy==1.24.3}'' specified in an onion comment (see Sec.~\ref{subsec:onion}). Since the user's existing version of the package satisfies this requirement, Davos will happily load it into the notebook. But if the user later upgrades their environment's \texttt{NumPy} version to v1.25.0 (perhaps as a result of installing a different package that depends on it) and subsequently reruns the notebook, the local version will longer satisfy the requirement, so Davos will install \texttt{NumPy} v1.24.3 into the notebook's project directory and load that version instead. From then on, further changes to the user's \texttt{Numpy} installation would not have any effect on Davos's behavior in this particular notebook, as a satisfactory version now exists in its project directory (if the version in the onion comment were changed, Davos would install it in the project directory in place of the current version). For efficiency, Davos projects will generally avoid duplicating dependencies already satisfied by the user's Python environment. However, if desired, adding \texttt{pip}'s \texttt{--ignore-installed} flag to an onion comment in the notebook will cause the smuggled package to be installed in the project directory whether or not it already exists locally.

it is possible to achieve a more ``traditional'' virtual environment-like experience by

In general, Davos will not duplicate dependencies satisfied by the existing Python environment in the project directory unless

.

.

For example, suppose a user runs a Davos-enhanced notebook that smuggles \texttt{NumPy}~\cite{HarrEtal20} with \texttt{numpy==1.24.3} specified in an onion comment (see Sec.~\ref{subsec:onion}), and happens to have \texttt{NumPy} v1.24.3 installed in their current Python environment.

.

.

.

For example, suppose a user runs a Davos-enhanced notebook that smuggles \texttt{NumPy}~\cite{HarrEtal20} v1.24.3 (specified via an onion comment; see Sec.~\ref{subsec:onion}) and happens to have that exact version of \texttt{NumPy} installed in their current Python environment.

.

.

.

Davos will load the user's existing \texttt{NumPy} version rather than

.

.

.

For example, suppose a user has \texttt{NumPy}~\cite{HarrEtal20} v1.24.3 installed in their current Python environment, and runs a Davos-enhanced notebook that smuggles that exact version of \texttt{NumPy} (specified via an onion comment; see Sec.~\ref{subsec:onion}). Since this requirement

%continue to be available and properly loaded, even if the user's Python environment changes. For example, if a user

%But in all cases, Davos's project system ensures that all dependencies of a given notebook are satisfied, and the user's Python environment is unaffected.
%But in all cases, Davos's project system ensures that the \textit{potential} effects of \texttt{smuggle} statements are confined to a single, isolated directory, and that all dependencies of a notebook can be satisfied while the user's Python environment remains unchanged. Additionally, because \texttt{smuggle} statements are evaluated each time the notebook is run, this system also ensures that all of its dependencies remain satisfied, even if the user's Python environment is changed. For example,

.

.

%But in all cases, Davos's project system ensures that the possible effects of \texttt{smuggle} statements are confined to a single, isolated directory, and that the user's Python environment remains unchanged. 
Additionally, because \texttt{smuggle} statements in a given notebook are evaluated each time that notebook is run, this system also ensures that all specified dependencies \textit{remain} satisfied, even if the user's main environment changes.

But in all cases, Davos's project system ensures that any packages installed are confined to an isolated directory, and that the user's Python environment is not affected. Additionally, because \texttt{smuggle} statements in a given notebook are evaluated each time that notebook is run, this system also ensures that all of its specified dependencies \textit{remain} satisfied, even if the user's main environment changes. For example, if a user

Additionally, because \texttt{smuggle} statements are evaluated each time a notebook is run, this system ensures that all required packages and specified package versions are always available and used, even if the notebook's main runtime environment is changed. For example, if a user

@@

%Notebook-specific projects are named for the absolute path to the notebook file.

 %Davos projects function similarly to simplified versions of standard Python virtual environments (e.g., created with the standard library's \texttt{venv} module or a third-party tool like \texttt{virtualenv}~\cite{BickEtal07}) with a few differences: they do not need to be manually activated and deactivated, they do not contain separate Python or \texttt{pip} executables, and they \textit{extend} the main Python environment rather than replace it.

>>>>

Thus, rather than constructing fully separate Python environments from scratch, Davos projects work by supplementing the user's existing environment with any additional packages (or specific package versions) needed to satisfy the dependencies of their corresponding notebooks.
In some cases, this might include every package smuggled into a notebook (e.g., if the notebook is run inside a freshly created, empty virtual environment). 
In other cases, the user's environment may already contain all required packages, and the notebook's project directory will be unused (in which case it will be deleted automatically when the notebook kernel is shut down). 
But regardless of the extent to which the existing environment is augmented, Davos's project system ensures that all smuggled packages are installed locally and can be loaded successfully at runtime, but that the contents of the user's Python environment are never altered. 
Additionally, because \texttt{smuggle} statements in a given notebook are evaluated every time it is run, this system also ensures that the notebook's requirements will continue to be satisfied even if the user's Python environment changes. 
For example, suppose a user has \texttt{NumPy}~\cite{HarrEtal20} v1.24.3 installed in their current Python environment and runs a Davos-enhanced notebook that smuggles \texttt{NumPy} with ``\texttt{numpy==1.24.3}'' specified in an onion comment (see Sec.~\ref{subsec:onion}). 
Since the user's existing version of the package satisfies this requirement, Davos will happily load it into the notebook. 
But if the user later upgrades their environment's \texttt{NumPy} version to v1.25.0 (perhaps as a result of installing a different package that depends on it) and subsequently reruns the notebook, the local version will longer satisfy the requirement, so Davos will install \texttt{NumPy} v1.24.3 into the notebook's project directory and load that version instead. 
From then on, further changes to the user's \texttt{Numpy} installation would not have any effect on Davos's behavior in this particular notebook, as a satisfactory version now exists in its project directory (if the version in the onion comment were changed, Davos would install it in the project directory in place of the current version). 
For efficiency, Davos projects will generally avoid duplicating dependencies already satisfied by the user's Python environment. 
However, if desired, adding \texttt{pip}'s \texttt{--ignore-installed} flag to an onion comment in the notebook will cause the smuggled package to be installed in the project directory whether or not it already exists locally.

@@
